--- 
title: "The EMaC R Manual"
author: "Alex Sciuto"
date: "Last Updated Febuary 2020"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
github-repo: rstudio/bookdown-demo
description: "This is a minimal example of using the bookdown package to write a book. The output format for this example is bookdown::gitbook."
---
--- 
title: "The EMaC R Manual"
author: "Alex Sciuto"
date: "Last Updated Febuary 2020"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
github-repo: rstudio/bookdown-demo
description: "This is a minimal example of using the bookdown package to write a book. The output format for this example is bookdown::gitbook."
---

# How to Install R {#download}

This is the EMaC manual for how to wrangle, analyze, and visualize data in **R**. I will review a few key packages, in addition to explaing their key functions, with real data examples. 

There are two things you need to do to install R on your computer. First, you would need to install the latest version of R which you would download and install from this link: [R-download](https://cloud.r-project.org/). Once you run and install R onto your computer, you would need to install R-Studio. R studio is the graphic user interface (GUI) where you can place all of your R-code. Follow this link: [R-Studio Desktop Download](https://rstudio.com/products/rstudio/download/). Once you have these two programs installed, all you need to do is launch R-studio Desktop and you are ready to go.



<!--chapter:end:index.Rmd-->


# Introduction to Programming in R {#intro}

Placeholder


## Variables in R
### Numeric
### Character
### Logical 
### Vector 
## Logical Operators & if-Statements

<!--chapter:end:01-intro.Rmd-->

# Introduction to Data Wrangling

To introduce data wrangling, we will be working with a dataset from a study conducted in our lab. Here are the details of the study in order to understand the data manipulation more: 


## The study

On each trial in this experiment (n = 174), each participant (n = 84) sees a target word very briefly (e.g., word) and then is prompted to select which of two letters was in a particular position of that word (e.g., _ _ _ ↓) - one letter was in the presented word (e.g., D) and the other letter is in an orthographic neighbor of the word (e.g., K). This is a 2 (visual field) x 3 (sentence context) experimental design: the target word is either presented in the fovea (i.e., center of the screen) or the parafovea (i.e., 3 degrees to the right of fixation). Prior to the target word, they see a sentence context presented via Rapid Serial Visual Presentation (RSVP) that constrains to target (i.e., makes the presented word predictable and the orthographic neighbor implausible), constrains to the alternative (i.e., makes the orthographic neighbor predictable and the presented word implausible), or is neutral (i.e., makes neither word predictable but both of them plausible). In addition, we collect data about the individual subjects’ language ability (i.e., their z-score on some test relative to the other participants), including spelling recognition (i.e., circle which words are spelled wrong), spelling dictation (i.e., write out words that they hear said), and phonological decoding (i.e., read aloud a list of words and a list of nonwords), and information about the lexical properties of the words (e.g., word frequency, cloze probability, orthographic neighborhood size, phonological neighborhood size, clustering coefficient, orthographic similarity to other words, which of the two words is higher frequency).

## Loading in Data

### Loading Packages

Before we start wrangling the data, we need to load in the packages we are using. In this lab, most of the data wrangling tools we will be using will be located in the `tidyverse()` package. What is a package? It is simply a group of functions that group of developers made that makes computations easier. To learn more about the `tidyverse()` package, you may click on this [link](https://www.tidyverse.org/) and read on it! For the data wrangling we will be doing in this section, we will be primarily using a package within tidyverse called `dplyr()`. to load in the package, simply type in the following code. What this code is saying is IF you don't have the `tidyverse()` installed, THEN install the package. After that, load the `tidyverse()` with the `library()` function.

```{r eval = FALSE, message = FALSE}
if(!require(tidyverse)){
    install.packages("tidyverse")
}
library(tidyverse) 
```

### setting the working directory

Now that we have the packages that we need, we need to load up our directory. What is a directory? It is essentially the folder we will be using to read in files, or export files into. In this directory, we have our data set of interest, `Topgown_Data`

```{r eval = FALSE, message = FALSE}
setwd("C:/Users/alexs/Documents/Github/Emac_Rmanual/TopGown/")
```

### Reading in Data

Now that we have our working directory set we can use the `read_csv()` function which requires one minimum argument to work which is the name of your `.csv` file. Here we are setting the variable `mydata` to the contents within the `.csv` file. It stores the contents of `Topgown_Data.csv` in what is called a data frame. What is a dataframe? it is simply a matrix where each row constitutes a variable (which can be any type you want), and rows are observations.

```{r}
mydata = read.csv("Topgown_Data.csv")
```

```{r echo = FALSE}
knitr::kable(mydata) %>%
  kable_styling("striped", full_width = F) %>%
  scroll_box(height = "200px")
```

<!--chapter:end:02-literature.Rmd-->

# Methods

We describe our methods in this chapter.

<!--chapter:end:03-method.Rmd-->

# Applications

Some _significant_ applications are demonstrated in this chapter.

## Example one

## Example two

<!--chapter:end:04-application.Rmd-->

# Final Words

We have finished a nice book.

<!--chapter:end:05-summary.Rmd-->

`r if (knitr:::is_html_output()) '
# References {-}
'`

<!--chapter:end:06-references.Rmd-->

