--- 
title: "The EMaC R Manual"
author: "Alex Sciuto"
date: "Last Updated Febuary 2020"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
github-repo: rstudio/bookdown-demo
description: "This is a minimal example of using the bookdown package to write a book. The output format for this example is bookdown::gitbook."
---

# How to Install R

This is the EMaC manual for how to wrangle, analyze, and visualize data in **R**. I will review a few key packages, in addition to explaing their key functions, with real data examples. 

There are two things you need to do to install R on your computer. First, you would need to install the latest version of R which you would download and install from this link: [R-download](https://cloud.r-project.org/). Once you run and install R onto your computer, you would need to install R-Studio. R studio is the graphic user interface (GUI) where you can place all of your R-code. Follow this link: [R-Studio Desktop Download](https://rstudio.com/products/rstudio/download/). Once you have these two programs installed, all you need to do is launch R-studio Desktop and you are ready to go.

The **bookdown** package can be installed from CRAN or Github:

```{r eval=FALSE}
install.packages("bookdown")
library(bookdown)
# or the development version
# devtools::install_github("rstudio/bookdown")
```

Remember each Rmd file contains one and only one chapter, and a chapter is defined by the first-level heading `#`.

To compile this example to PDF, you need XeLaTeX. You are recommended to install TinyTeX (which includes XeLaTeX): <https://yihui.name/tinytex/>.

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```
